R Notebook

### 2017-04-03 doc/RMD/MLG-distribution.Rmd
massive update to plots. These looks quite good.

```{r}
library('tidyverse')
library('assertr')
library('poppr')
library('igraph')
library("ggraph")
```

@@

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
dat
```

@@

```{r}
setPop(dat) <- ~Region
dat
```

@@

```{r}
datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
adjmat  <- datmlg %*% t(datmlg) 
cols    <- sort(colnames(adjmat))
adjmat  <- adjmat[cols, cols]
adjmat
crosses
```

@@

```{r, fig.width = 10, fig.height = 10}
g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
V(g)$size   <- diag(adjmat)
g           <- delete_vertices(g, degree(g) == 0)
shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
V(g)$weight <- 1 - shared_mlg/V(g)$size
el          <- as_adj_edge_list(g)
el          <- el[lengths(el) > 0]
popgraphs <- setNames(vector(mode = "list", length = length(el)), names(el))
for (v in names(el)){
  idx  <- el[[v]]
  mlgs <- crosses %>%           # How to get all MLGs from a single population:
    filter(Population == v) %>%         # Grab only the population e and then
    select(MLG) %>%                     # remove everything but the MLGs to do an
    inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
    filter(Population != v) %>%         # remove the query population to give
    arrange(Population)                 # the neigboring populations in order.
  MLGS <- as.character(mlgs$MLG)
  E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
  E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
  popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
}
par(mfrow = c(3, 4))
for (i in names(popgraphs)){
  pg <- popgraphs[[i]]
  labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
  labs <- ifelse(duplicated(labs), NA, labs)
  plot(pg, 
       main = i, 
       layout = layout_as_star(pg, center = i), 
       edge.width = E(pg)$weight,
       edge.label = labs)
}
par(mfrow = c(1, 1))
center_node <- degree(g) %>% which.max() %>% names()

new_layout <- structure(c(1, 0.809016994374947, 0.309016994374947, -0.309016994374947, 
-0.809016994374947, -1, -0.809016994374947, -0.309016994374948, 
0, 0.309016994374947, 0.809016994374947, 0, 0.587785252292473, 
0.951056516295154, 0.951056516295154, 0.587785252292473, 1.22464679914735e-16, 
-0.587785252292473, -0.951056516295154, 0, -0.951056516295154, 
-0.587785252292473), .Dim = c(11L, 2L), .Dimnames = list(c("MN", 
"AU", "OR", "FR", "WA", "CA", "NE", "CO", "MI", "ND", "NY"), 
    c("x", "y")))
new_layout <- new_layout[V(g)$name, ]
g <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
plot(g, 
     layout = rbind(new_layout, new_layout),
     edge.width = E(g)$weight,
     edge.label = NA)

glay <- create_layout(g, "manual", node.positions = as.data.frame(rbind(new_layout, new_layout)))
ggraph(glay) +
  geom_edge_fan(aes(alpha = weight + 1)) +
  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = weight)) +
  geom_node_label(aes(label = name), repel = TRUE) +
  viridis::scale_fill_viridis(option = "C") +
  coord_fixed() +
  theme_void() +
  labs(list(
    title = "Shared haplotypes across regions",
    fill = "Number of\nGenotypes",
    alpha = "Fraction of\nprivate genotpes",
    edge_alpha = "Populations\nper haploytpe",
    caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
  ))
```

>>>>

```{r}
library('tidyverse')
library('assertr')
library('poppr')
library('igraph')
library("ggraph")
library('magrittr')
```

@@

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
dat

repeat_lengths <-
  c(
  `5-2` = 2.000000,
  `5-3` = 0.500000,
  `6-2` = 6.000000,
  `7-2` = 2.000000,
  `8-3` = 2.000000,
  `9-2` = 2.000000,
  `12-2` = 2.000000,
  `17-3` = 3.000000,
  `20-3` = 2.000000,
  `36-4` = 0.500000,
  `50-4` = 0.500000,
  `55-4` = 4.000000,
  `92-4` = 2.000000,
  `106-4` = 4.000000,
  `110-4` = 4.000000,
  `114-4` = 4.000000
  )
repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
```

```{r excel_data}
ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
  select(-1) %>%                # removing first column, which is empty
  gather(locus, allele, -1) %>% # gather all loci into tidy columns
  mutate(locus = trimws(locus)) %>% # remove (F) designator
  mutate(allele = as.integer(allele)) %>% # force alleles to integers
  spread(locus, allele) %>%
  slice(-n())
ex <- ex[!names(ex) %in% locNames(dat)]

# Function to select an adjacent allele. It will select the
# next allele if the next allele is not missing and it's distance
# is one away and the previous allele for the same conditions.
# If none of the conditions are met, it will retain the allele.
cromulent_allele <- Vectorize(function(lower, allele, higher){
  if (!is.na(higher) && abs(allele - higher) == 1){
    out <- higher
  } else if (!is.na(lower) && abs(allele - lower) == 1){
    out <- lower
  } else {
    out <- allele
  }
  out
})
ex
exsummary <- ex %>% 
  gather(locus, allele, -1) %>% # tidy the data
  group_by(locus, allele) %>%   
  summarize(n = n()) %>%        # summarize by count 
  ungroup() %>%
  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
  select(locus, new_allele, allele)
exsummary
corrected_loci <- ex %>% gather(locus, allele, -1) %>%
  left_join(exsummary, by = c("locus", "allele")) %>%
  mutate(allele = new_allele) %>%
  select(-new_allele) %>%
  spread(locus, allele)
datdf <- genind2df(dat, usepop = FALSE) %>% bind_cols(corrected_loci[-1])
dat  <- df2genind(datdf, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% as.genclone()
```

@@

```{r}
setPop(dat) <- ~Region
dat
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
genotype_curve(dat, sample = 1000)
locus_table(dat)
# rl <- fix_replen(dat, repeat_lengths)
# mlg.filter(dat, distance = bruvo.dist, replen = rl) <- .Machine$double.eps
dat
genotype_curve(dat[loc = keeploci, mlg.reset = TRUE])
# dat <- dat[loc = keeploci, mlg.reset = TRUE]
dat
```

@@

```{r}
make_graph_list <- function(dat){ # dat is a genclone/snpclone object
  datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
  crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
  adjmat  <- datmlg %*% t(datmlg) 
  cols    <- sort(colnames(adjmat))
  adjmat  <- adjmat[cols, cols]
  # Creating Graph
  g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
  V(g)$size   <- diag(adjmat)
  g           <- delete_vertices(g, degree(g) == 0)
  shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
  V(g)$weight <- 1 - shared_mlg/V(g)$size
  el          <- as_adj_edge_list(g)
  el          <- el[lengths(el) > 0]
  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
  for (v in names(el)){
    idx  <- el[[v]]
    mlgs <- crosses %>%           # How to get all MLGs from a single population:
      filter(Population == v) %>%         # Grab only the population e and then
      select(MLG) %>%                     # remove everything but the MLGs to do an
      inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
      filter(Population != v) %>%         # remove the query population to give
      arrange(Population)                 # the neigboring populations in order.
    MLGS <- as.character(mlgs$MLG)
    E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
    E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
    popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
  }
  popgraphs[["total"]] <- g
  popgraphs
}

plot_mlg_graph <- function(g, glayout = NULL){
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- (1 - V(g)$weight) * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  ggraph(glay) +
    geom_edge_fan(aes(alpha = weight + 1), width = 0.75) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
    geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) +
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(0.25, 1), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nGenotypes",
      edge_alpha = "Populations\nper haploytpe",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

plot_mlg_subgraph <- function(graphlist){
  for (i in names(graphlist)){
    pg   <- graphlist[[i]]
    labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
    labs <- ifelse(duplicated(labs), NA, labs)
    plot(pg, 
         main = i, 
         layout = layout_as_star(pg, center = i), 
         edge.width = E(pg)$weight,
         edge.label = labs)
  }
}

```

@@

```{r, fig.width = 7, fig.height = 7}
graph16loc <- make_graph_list(dat)
graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])

good_layout <- read.table(
text = 
"   x                            y
AU  0.8090169943749470071737  5.877852522924730260812e-01
CA -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
FR -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
OR  0.3090169943749470071737  9.510565162951539752711e-01
WA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()

par(mfrow = c(3, 4))
plot_mlg_subgraph(graph16loc[-length(graph16loc)])

par(mfrow = c(3, 4))
plot_mlg_subgraph(graph11loc[-length(graph11loc)])

par(mfrow = c(1, 1))
plot_mlg_graph(graph16loc$total, good_layout)
plot_mlg_graph(graph11loc$total, good_layout)
```

```{r, eval = FALSE}
null_vertices <- length(V(g)) %>% seq(from = ./2 + 1, to = ., by = 1)
sg <- g %>% 
  delete_vertices(null_vertices) %>%
  cluster_spinglass()
modularity(sg)
communities(sg)
```
